---
title: "02-demographic-data"
author: "Sasha Darknell-Brown"
date: '2023-10-17'
output: html_document
---

# Set up R Markdown document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../output/02-demographic-data/")) #Set root directory
```

```{r}
getwd()
```

# START YOUR CHANGES HERE

#Load packages
```{r}
install.packages("tidyverse") #Download tidyverse package and run it 
library(tidyverse)
```


```{r}
install.packages("readr") #(not needed, but it imports the package to download your datasets)
library(readr)
```


#Import data sets 
```{r}
my_data <- read_csv("../../metadata/metabolomics-sample-list.csv") #brings the data set in 
```

```{r}
my_data <- read_csv("../../data/time-to-right.csv") #load the time to right data 
library()
```



```{r}
names(time_to_right) #Shows me the names of the columns in the dataset
```
```{r}
names(metabolomics_sample_list) #Shows names of columns 
```


```{r}
my_data <- read_csv("../../data/crab-metadata.csv") #load the time to right data
```


```{r}
unique(crab_metadata$treatment)
```



```{r}
levels(crab_metadata$treatment)
```


#Making Graph 1 

```{r}
crab_metadata$treatment <- factor(
  crab_metadata$treatment, 
  levels = c("5C", "15C", "25C")
)
```


```{r}
crab_metadata %>%
  ggplot(aes(x = (treatment),
             y = `carapace-width`,
             color = treatment,
             shape = treatment)) +
  geom_boxplot() +
  geom_jitter(position = position_dodge(width = 0.90)) +
  labs(x = "Temperature (ºC)", y = "Carapace Width (mm)") +
  scale_color_manual(values = c("green", "purple", "lightblue")) +
  scale_shape_manual(values = c(15, 16, 17)) +
  theme_classic() 
 
```


```{r}
str(crab_metadata$treatment) #checking the treatment variable to see if it was changed correctly
table(crab_metadata$treatment, useNA = "always")
```


#Graph 1: Joining the two datasets (TTR & crab_metadata)



```{r}
names(crab_metadata)
```


```{r}
names(time_to_right)
```


```{r}
crab_metadata_day <- left_join(
  crab_metadata,
  time_to_right,
  by = c("ID" = "crab-ID")
)
```

```{r}
names(crab_metadata_day)
```

```{r}
unique(crab_metadata_day$date) #checking to see if the merge worked
```





#Reference for ggplot 
```{r}
metabolomics_sample_list %>% #Uses this data frame for the graph 
  ggplot(., aes(x = as.character(timepoint), #tells us what column we are pulling data from to make the x-axis 
                y= avg_righting, #tells us what column we are taking data from to use for Y 
                color = population, #changes the color for each population of crab 
                shape = population)) + #Puts different shapes for the two populations 
  geom_boxplot() + #makes a boxplot
  geom_jitter(position = position_dodge(width = 0.90)) + #spaces out the data points to keep them from overlapping
  labs(x = "Timepoint (Hours)", y = "Average Righting Response (s)") + #Allows us to name the X and Y axes 
  scale_color_manual(values = c("green", "purple"), #Allows us to pick the colors for out two crab populations
                     name = "Population", #Legend title 
                     labels = c("Maine", "Washington")) + #Names of the populations for the legend
  scale_shape_manual(values = c(15, 16), #Lets us pick the shapes we want for the data points 
                     name = "Population", #name of the legend --> putting it here as well means we will only get 1 legend rather than 2
                     labels = c("Maine", "Washington")) + #what to name the two populations in the legend
  theme_classic() + #gets rid of the grey background in the graph 
  theme(legend.position = "bottom") #Allows us to re-position the legend 
```










# DON'T EDIT BELOW HERE

# Survival analysis

I want to determine if there was a significant difference in survival time between temperature treatments.

## Import data

```{r}
survivalData <- read.csv("../../data/survivorship.csv", header = TRUE) %>%
  select(-date) %>%
  mutate(., day = day - 5) %>%
  filter(., day >= 0 ) %>%
  column_to_rownames(., var = "day") %>%
  t(.) %>% 
  as.data.frame(.) %>%
  rownames_to_column(., var = "tank") #Import data frame and remove date column. Convert day into actual days of the treatment part of the experiment by subtracting 5. Remove acclimation survival counts. Convert the day column into rownames, transpose, and convert into a dataframe.
head(survivalData)
```

## Format data

I need to convert `survivalData` into a dataframe with three columns: time (days until death), status (1 = event, 0 = censored), treatment (15C = ambient, 5C = cold, 25C = hot).

```{r}
mortalityData <- survivalData[NA,] %>%
  mutate(., tank = survivalData$tank) %>%
  mutate(., "0" = rep(0, times = 9)) %>%
  mutate(., "50" = survivalData$`49`) #Create an empty dataframe with the same structure as survivalData. First day had 0 mortalities. Add a column with remaining live crabs as day 50, then add tank IDs
rownames(mortalityData) <- rownames(survivalData) #Use the same row names
for (i in 3:51) {
  mortalityData[,i] <- survivalData[,(i-1)] - survivalData[,i]
} #For day 1 through 49, subtract the number of alive crabs each day from how many there were the day before to calculate how many died each day.
head(mortalityData, n = 9) #Confirm dataframe creation
```

```{r}
mortalityData <- mortalityData %>%
  mutate(., "42" = c(0, 0, 0, 1, 0, 0, 0, 0, 0)) %>%
  mutate(., "42c" = c(2, 9, 9, 2, 8, 6, 2, 8, 6)) %>%
  select(., c(1:44, 53, 45:52)) #On day 42, 2 crabs were censored from each cold tank, but an additional crab died in T4. All remaining crabs from the control and warm tanks were censored. First, modify day 42 data to only include mortality, then add information of censored crabs in a new column. Rearrange columns using select.
head(mortalityData, n = 9) #Confirm changes
```

Samples were collected (censored == 0) on day 1 (all), day 6 (15ºC and 25ºC), day 8 (5ºC), day 26 (all), 42 (all), and 50 (5ºC)

```{r}
mortalityDataLong <- mortalityData %>%
  pivot_longer(., cols = -tank, names_to = "day", values_to = "num_died") %>%
  mutate(status = case_when(day == "1" | day == "26" | day == "42c" | day == "50" ~ 0,
                            day == "6" & tank == "T2" | day == "6" & tank == "T5" | day == "6" & tank == "T8" ~ 0,
                            day == "6" & tank == "T3" | day == "6" & tank == "T6" | day == "6" & tank == "T9" ~ 0,
                            day == "8" & tank == "T1" | day == "8" & tank == "T4" | day == "8" & tank == "T7" ~ 0)) %>%
  mutate(status = case_when(status == 0 ~ 0,
                            is.na(status) == TRUE ~ 1)) %>%
  filter(., is.na(num_died) == FALSE) %>%
  filter(., num_died != 0) %>%
  mutate(., day = replace(day, day == "42c", 42)) %>%
  mutate(., day = replace(day, day == "50", 49)) %>%
  mutate(time = as.numeric(day)) %>%
  select(., -day) %>%
  uncount(., weights = num_died) %>%
  mutate(., treatment = case_when(tank == "T1" | tank == "T4" | tank == "T7" ~ "5C",
                                  tank == "T2" | tank == "T5" | tank == "T8" ~ "13C",
                                  tank == "T3" | tank == "T6" | tank == "T9" ~ "30C")) %>%
  select(., -tank)
#Pivot the dataframe longer, using former column names as experimental day indication and values as number of crabs that died each day. Use case_when to assign status information (0 = censored/sampled, 1 = mortality) based on day and tank information. Remove rows where the num_died == NA (aka for 15ºC and 25ºC tanks that ended early) and for rows where num_died == 0. Replace "42c" with 42 and 50 with 49, since those distinctions were used to easily separate censored and uncensored crabs. Create a column "time" which is a numeric version of day, then remove day. Create a duplicate row for each crab that died using uncount. Use case_when to assign treatment information to each tank and remove the tank column. 
head(mortalityDataLong)
```

## Compare groups

This analysis is based on [the following tutorial from Dr. Zabor](https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html) and a similar analysis of Olympia oyster larval survival in different temperature treatments by Dr. Lindsay Alma.

### Estimate median survival time

```{r}
mortalityDataLong %>%
  survfit(Surv(time, status, type = "right") ~ treatment, data = .) #Use survfit to estimate median survival time. There is not enough data to estimate median survival. Not a useful analysis given the low sample size and low mortality data.
```

### Compare survival times between groups

```{r}
mortalityDataLong %>%
  survdiff(Surv(time, status, type = "right") ~ treatment, data = .) #Use a log-rank test to compare temperature treatments. No significant effect of treatment on mortality (Chisq= 3  on 2 degrees of freedom, p= 0.2 )
```

According to L. Alma, "When survival curves cross, the log-rank test should not be used because the probability of a type II error will be increased, and the test has low power." I'll also use a Cox regression model to understand differences in survival by treatment.

```{r}
coxModel <- mortalityDataLong %>%
  coxph(Surv(time, status, type = "right") ~ treatment, data = .) #Use a Cox regression model to investigate survival differences by treatment. 
coxModel #No significant difference between treatments

#                 coef exp(coef) se(coef)      z     p
# treatment30C  0.7787    2.1787   0.7072  1.101 0.271
# treatment5C  -0.4520    0.6363   0.9129 -0.495 0.621
# 
# Likelihood ratio test=2.87  on 2 df, p=0.2385
# n= 137, number of events= 11 

summary(coxModel) #Confidence interval information for each treatment as compared to the base condition (15)

#              exp(coef) exp(-coef) lower .95 upper .95
# treatment30C    2.1787      0.459    0.5448     8.713
# treatment5C     0.6363      1.571    0.1063     3.809
```

The `exp(coef)` parameter is the hazards ratio. The HR means that each additional day is associated with a HR-fold increase in chance of death. A HR > 1 means an increase in the hazard when compared to the control. There is an 2.18-fold increase in death as the experiment progresses for the 25ºC treatment, while only a 0.63 increase in death for the 5ºC treatment. However, these are not significantly different than what is expected for the 15ºC treatment.

### Plot Kaplan-Meier curves

```{r}
mortalityDataLong %>%
  survfit2(Surv(time, status, type = "right") ~ treatment, data = .) %>%
  ggsurvfit(linewidth = 1, linetype_aes = TRUE) +
  scale_color_manual(values = c(plotColors[2], plotColors[1], plotColors[3]),
                     labels = c("15", "25", "5"),
                     name = "Temperature (ºC)") +
  scale_linetype_manual(values = c(1, 6, 4),
                        labels = c("15", "25", "5"),
                        name = "Temperature (ºC)") +
  scale_x_continuous(breaks = c(seq(0, 45, 5), 49)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  labs(x = "Days", y = "Overall Survival Probality") + 
  annotate(geom = "text", x = -1, y = 0.2, 
           label = expression(paste(bolditalic("Overall: "), Chi^2, " = 3.00, p = 0.20")),
           hjust = 0) +
  annotate(geom = "text", x = -1, y = 0.15, 
           label = expression(paste(bolditalic("15ºC vs. 25ºC: "), "HR = 2.18, p = 0.27")),
           hjust = 0) +
  annotate(geom = "text", x = -1, y = 0.1, 
           label = expression(paste(bolditalic("15ºC vs. 25ºC: "), "HR = 0.63, p = 0.62")),
           hjust = 0) +
  theme_classic(base_size = 15) #Plot KM curves based on the Surv object analyzing survival time by treatment, with "right" indicating the kind of censoring (samples still alive at the end of the experiment were "right" censored). Customize line width, line type, colors, legend, axes, and labels. Add annotations for statistics. Use default R plotting theme with a base font size of 15.
ggsave("figures/KM-curves.pdf", height = 8.5, width = 11)
```

# Genotype characterization

```{r}
genotypeData <- read.csv("../../data/genotype/2023-crab-genotype.csv", header = TRUE, na.strings = "") %>%
  filter(., is.na(final.genotype) == FALSE) %>%
  select(., crab.ID, treatment, final.genotype) #Import genotype data. Remove rows without final genotypes
head(genotypeData) #Confirm genotype dataframe creation
```

```{r}
genotypeData %>%
  group_by(., final.genotype) %>%
  summarize(., count = n())
```

I took the above numbers and put them into [this Hardy-Weinberg equilibrium calculator](https://www.cog-genomics.org/software/stats).

P-value: 0.4238149514542

This population is in Hardy-Weinberg equilibrium!

```{r}
genotypeData %>%
  group_by(., final.genotype) %>%
  summarise(., count = n()) %>%
  ggplot(., aes(x= "", y = count, fill = final.genotype)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  # geom_text(aes(label = paste0(round((count/sum(count)*100), digits = 0), "%")), position = position_stack(vjust = 0.5), size = 6) +
  scale_fill_manual(name = "Genotype",
                    values = c("grey90", "grey45", "grey0")) +
  theme_void(base_size = 15) #Group genotype data by genotype, then count the number of crabs in each genotype. Create a pie chart for genotype distribution, irrespective of treatment by using geom_bar then setting polar coordinates. Add text labels for % crabs in each genotype. Create a manual color scale.
ggsave("figures/genotype.pdf", height = 8.5, width = 11)
```

```{r}
facet_labels <- c("15C" = "15ºC",
                  "25C" = "25ºC",
                  "5C" = "5ºC") #Assign labels for facets
```

```{r}
genotypeData %>%
  group_by(., treatment, final.genotype) %>%
  summarise(., count = n()) %>%
  ggplot(., aes(x= "", y = count, fill = final.genotype)) +
  geom_bar(stat = "identity", width = 1, position = "fill") +
  coord_polar("y", start = 0) +
  facet_wrap(vars(treatment), ncol = 3, labeller = labeller(treatment = facet_labels)) +
  scale_fill_manual(name = "Genotype",
                    values = c("#BAE4B3", "#74C476", "#006D2C")) +
  theme_void(base_size = 15)  + theme(strip.text.x = element_text(size = 15)) #Group genotype data by genotype, then count the number of crabs in each genotype. Create a pie chart for genotype distribution, irrespective of treatment by using geom_bar then setting polar coordinates. Create a manual color scale and adjust text of facet labels.
ggsave("figures/genotype-by-treatment.pdf", height = 8.5, width = 11)
```
                                      
                                      