---
title: "06g-gene-expression-edgeR"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#loading required packages
```{r}
require(tidyverse)
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("DESeq2")
BiocManager::install("edgeR")
library(DESeq2)
library(ggplot2)
library(edgeR)
require("RColorBrewer")
install.packages("pheatmap")
require(pheatmap)
```
# link for the dataset download
```{r}
# https://osf.io/ea8mq/files/xnt56
```

# MAC import dataset
```{r}
require(tidyverse)
salmon.data <- read.delim("../../../Downloads/salmon.genes.counts.matrix", header=TRUE , check.names = FALSE)
```
# WINDOWS import dataset
```{r}
require(tidyverse)
salmon.data <- read.delim("../../../../Downloads/salmon.genes.counts.matrix", header=TRUE , check.names = FALSE)
```
# alter salmon.data so ID column doesn't mess with data
```{r}
colnames(salmon.data)[1] <- "ID"
head(salmon.data)
```



# beginning EdgeR analysis by making the table
```{r}
counts <- salmon.data
colnames(counts) <- counts[1, ]
dgList <- DGEList(counts=counts, genes=rownames(counts))
dgList
head(dgList$counts) #Many rows!
head(dgList$genes)
```
# finding counts per million expression levels
```{r}
countsPerMillion <- cpm(dgList)
summary(countsPerMillion)
```
# removing expression levels that are too low
```{r}
countCheck <- countsPerMillion > 1
head(countCheck)
keep <- which(rowSums(countCheck) >= 2)
dgList <- dgList[keep,]
summary(cpm(dgList)) 
```
# beginning the analysis
```{r}
?calcNormFactors
dgList <- calcNormFactors(dgList, method="TMM")
```
```{r}
plotMDS(dgList)
```
# comparing different treatment groups to each other
```{r}
sampleType<- rep("15", ncol(dgList)) #N=normal; T=tumour
sampleType[grep("05", colnames(dgList))] <- "05"
#'grep' is a string matching function.
sampleReplicate <- paste("S", rep(1:3, each=2), sep="")
designMat <- model.matrix(~sampleReplicate + sampleType)
designMat
```

```{r}
dgList <- estimateGLMCommonDisp(dgList, design=designMat)
dgList <- estimateGLMTrendedDisp(dgList, design=designMat)
dgList <- estimateGLMTagwiseDisp(dgList, design=designMat)
```
# Create a DGEList object
```{r}
dge <- DGEList(counts = salmon.data, group = group_factor)
keep <- rowSums(cpm(dge) > 1) >= min(table(group_factor))
dge <- dge[keep, , keep.lib.sizes = FALSE]
```



# Trying Method 2 - Gemini Assisted
# making dataset renaming columns so they fit DGEList
```{r}
counts_data <- salmon.data

sample_names <- colnames(counts_data)
raw_groups <- sapply(strsplit(sample_names, "-"), `[`, 1)
groups <- factor(paste0("G", raw_groups)) 
```

# making DGEList object

```{r}
y <- DGEList(counts = counts_data, group = groups)
```

# filtering and normalization

```{r}
keep <- filterByExpr(y)
table(keep) # 59,912 genes were kept
y <- y[keep, , keep.lib.sizes=FALSE]
y <- calcNormFactors(y)
```

# creating design matrix

```{r}
design <- model.matrix(~ 0 + groups)
colnames(design) <- levels(groups) 
```

# fitting model

```{r}
y <- estimateDisp(y, design, robust=TRUE)
fit <- glmQLFit(y, design, robust=TRUE)
```

# comparing 25C and 5C treatments

```{r}
my.contrasts <- makeContrasts(Group25_vs_5 = G25 - G5, levels = design)
qlf_25_vs_5 <- glmQLFTest(fit, contrast=my.contrasts[,"Group25_vs_5"])
topTags(qlf_25_vs_5, n=10) # change n= value to see more genes, or remove it to see all. Negative logFC means downregulation of first treatment vs second treatment
```

# comparing 15C and 5C treatments

```{r}
my.contrasts <- makeContrasts(Group15_vs_5 = G15 - G5, levels = design)
qlf_15_vs_5 <- glmQLFTest(fit, contrast=my.contrasts[,"Group15_vs_5"])
topTags(qlf_15_vs_5, n=10)
```

# comparing 15C and 25C treatments

```{r}
my.contrasts <- makeContrasts(Group15_vs_25 = G15 - G25, levels = design)
qlf_15_vs_25 <- glmQLFTest(fit, contrast=my.contrasts[,"Group15_vs_25"])
topTags(qlf_15_vs_25, n=10)
```