---
title: "06g-gene-expression-edgeR"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#loading required packages
```{r}
require(tidyverse)
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("DESeq2")
BiocManager::install("edgeR")
library(DESeq2)
library(ggplot2)
library(edgeR)
require("RColorBrewer")
install.packages("pheatmap")
require(pheatmap)
```
# link for the dataset download should not be chunk of code
```{r}
# https://osf.io/ea8mq/files/xnt56
```

# MAC import dataset
```{r}
require(tidyverse) #this loads the package into the session
salmon.data <- read.delim("../../../Downloads/salmon.genes.counts.matrix", #downloads data and names it "salmon.data". 
    header=TRUE , check.names = FALSE) # "header = TRUE" tells the system that the first row is not numerical. "check.names = FALSE" tells the system to leave the headers exactly how they are.
```
# WINDOWS import dataset
```{r}
require(tidyverse)
salmon.data <- read.delim("../../../../Downloads/salmon.genes.counts.matrix", header=TRUE , check.names = FALSE)
```
# alter salmon.data so ID column doesn't mess with data
```{r}
colnames(salmon.data)[1] <- "ID" #changing the first columns name to "ID"
head(salmon.data) #shows the first 6 rows of the dataset which allows you to see if your column name was changed accordingly 
```

# making dataset renaming columns so they fit DGEList
```{r}
counts_data <- salmon.data #Creates a copy of the data set and names it "countsdata"
rownames(counts_data) <- counts_data[, 1] #sets the row names equal to the values in the first column
counts_data <- counts_data[, -1] #Exclude column 1 
colnames(counts_data)[colnames(counts_data) == "25-168"] <- "15-168"
colnames(counts_data)[colnames(counts_data) == "5-009"] <- "15-009"
colnames(counts_data)[colnames(counts_data) == "25-053"] <- "15-053"
colnames(counts_data)[colnames(counts_data) == "25-119"] <- "15-119"
colnames(counts_data)[colnames(counts_data) == "5-074"] <- "15-074"
colnames(counts_data)[colnames(counts_data) == "5-122"] <- "15-122"
colnames(counts_data)[colnames(counts_data) == "5-182"] <- "15-182"
colnames(counts_data)[colnames(counts_data) == "5-189"] <- "15-189"
colnames(counts_data)[colnames(counts_data) == "25-051"] <- "15-051"
colnames(counts_data)[colnames(counts_data) == "25-110"] <- "15-110"
colnames(counts_data)[colnames(counts_data) == "25-202"] <- "15-202"
colnames(counts_data)[colnames(counts_data) == "5-133"] <- "15-133"
sample_names <- colnames(counts_data) #gets the column names from counts_data and stores them in a vector called sample_names
raw_groups <- sapply(strsplit(sample_names, "-"), `[`, 1) #splits the sample name wherever it finds a hyphen. "[, 1" says to take the first part of each split. "sapply" repeats the extraction for every sample
groups <- factor(paste0("G", raw_groups)) #"paste0("G", raw_groups)" joins text together without spaces. "G" is added in front of each number. Factor makes the labels "G#" into a category.
```

# making DGEList object

```{r}
y <- DGEList(counts = counts_data, group = groups) #function from the edgeR package that stores the count matrix, group labels, total reads per sample, normalization factors. "counts = counts_data" tells edgeR that they are the raw gene counts. "group = groups" tells edgeR which samples belong to which group. 
```

# filtering and normalization

```{r}
keep <- filterByExpr(y) #filters out genes with very low expression 
table(keep) # 59,912 genes were kept
y <- y[keep, , keep.lib.sizes=FALSE] #tells edgeR which genes to keep 
y <- calcNormFactors(y) #normalizes the dataset
```

# creating design matrix

```{r}
design <- model.matrix(~ 0 + groups)
colnames(design) <- levels(groups) 
```

# fitting model

```{r}
y <- estimateDisp(y, design, robust=TRUE)
fit <- glmQLFit(y, design, robust=TRUE)
```

# comparing 25C and 5C treatments

```{r}
my.contrasts <- makeContrasts(Group25_vs_5 = G25 - G5, levels = design)
qlf_25_vs_5 <- glmQLFTest(fit, contrast=my.contrasts[,"Group25_vs_5"])
topTags(qlf_25_vs_5, n=10)
results_25vs5 <- topTags(qlf_25_vs_5, n=Inf)$table
filtered_25vs5 <- subset(results_25vs5, FDR < 0.05)
nrow(filtered_25vs5) # 33,733 DEGs 
# change n= value to see more genes, or remove it to see all. Negative logFC means downregulation of first treatment vs second treatment
```

# comparing 15C and 5C treatments

```{r}
my.contrasts <- makeContrasts(Group15_vs_5 = G15 - G5, levels = design)
qlf_15_vs_5 <- glmQLFTest(fit, contrast=my.contrasts[,"Group15_vs_5"])
topTags(qlf_15_vs_5, n=10)
results_15vs5 <- topTags(qlf_15_vs_5, n=Inf)$table
filtered_15vs5 <- subset(results_15vs5, FDR < 0.05)
nrow(filtered_15vs5) # 8557 DEGs
```

# comparing 15C and 25C treatments

```{r}
my.contrasts <- makeContrasts(Group15_vs_25 = G15 - G25, levels = design)
qlf_15_vs_25 <- glmQLFTest(fit, contrast=my.contrasts[,"Group15_vs_25"])
topTags(qlf_15_vs_25, n=10)
results_15vs25 <- topTags(qlf_15_vs_25, n=Inf)$table
filtered_15vs25 <- subset(results_15vs25, FDR < 0.05)
nrow(filtered_15vs25) # 888 DEGs
```

# making the PCA
```{r}
logCPM <- cpm(y, log = TRUE)
pca_results <- prcomp(t(logCPM), scale. = TRUE)
pca_data <- data.frame(
  Sample = rownames(pca_results$x),
  PC1 = pca_results$x[, 1],
  PC2 = pca_results$x[, 2],
  Group = groups 
)

pca_var <- pca_results$sdev^2
pca_var_perc <- round(pca_var / sum(pca_var) * 100, 1)
```

# plotting the PCA 

```{r}
ggplot(pca_data, aes(PC1, PC2, group = Group, color = Group, shape = Group)) +
  geom_point(size=3) + 
  stat_ellipse() + #Add 95% confidence interval ellipses around data
  xlab(paste0("PC1: ",pca_var_perc[1],"% variance")) +
  ylab(paste0("PC2: ",pca_var_perc[2],"% variance")) + 
  coord_fixed() +
  scale_color_manual(values = c( "#969696", "#CB181D", "#2171B5"), 
                     name = "Temperature (ºC)",
                     labels = c("15", "25", "5")) +
  scale_shape_manual(values = c(17, 16, 15),
                     name = "Temperature (ºC)",
                     labels = c("15", "25", "5")) +
  theme_classic() + theme(legend.position = "top")

ggsave("../output/06f-gene-expression/EdgeR_PCAcraball.pdf", width = 11, height = 8.5)
```


