---
title: "06g-gene-expression-edgeR"
output: html_document
---

```{r}
getwd() #saves to your working directory for your plots (ggsave)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#loading required packages
```{r}
require(tidyverse)
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("DESeq2")
BiocManager::install("edgeR")
library(DESeq2)
library(ggplot2)
library(edgeR)
require("RColorBrewer")
install.packages("pheatmap")
require(pheatmap)
```
# link for the dataset download should not be chunk of code
 https://osf.io/ea8mq/files/xnt56


# MAC import dataset
```{r}
require(tidyverse) #this loads the package into the session. This package has the ggplot, dplyr, tidyr, etc...things that make data analysis easier.

salmon.data <- read.delim("../../wc-green-crab/metadata/salmon.genes.counts.matrix", #downloads data and names it "salmon.data". *make sure to change the directory to where it is on your computer.
                          
    header=TRUE , check.names = FALSE) # "header = TRUE" tells the system that the first row is not numerical. "check.names = FALSE" tells the system to leave the headers exactly how they are.
```

# alter salmon.data so ID column doesn't mess with data
```{r}
colnames(salmon.data)[1] <- "ID" #changing the first columns name to "ID"
head(salmon.data) #shows the first 6 rows of the dataset which allows you to see if your column name was changed accordingly 
```

# making dataset renaming columns so they fit DGEList
```{r}
counts_data <- salmon.data #Creates a copy of the data set and names it "countsdata"
rownames(counts_data) <- counts_data[, 1] #This brought all of the original rows with their titles, and also duplicated the first row, one with the "ID" title, and the original row without a title. 
```


```{r}
counts_data <- counts_data[, -1] #Exclude column 1 with the "ID" as the title. Kept the original row without a title name.
head(counts_data)
```

#Renaming Specific Columns in the Counts_Data
*colnames(counts_data) --> gathers the names of the columns in dataset
* == "25-168" --> checks the columns with this name 
* <- "15-168" --> changes the original name to this 
* The reason why the names needed to be changed is because the column names were not matching the names in the count matrix (salmon.data) 
```{r}
colnames(counts_data)[colnames(counts_data) == "25-168"] <- "15-168"
colnames(counts_data)[colnames(counts_data) == "5-009"] <- "15-009"
colnames(counts_data)[colnames(counts_data) == "25-053"] <- "15-053"
colnames(counts_data)[colnames(counts_data) == "25-119"] <- "15-119"
colnames(counts_data)[colnames(counts_data) == "5-074"] <- "15-074"
colnames(counts_data)[colnames(counts_data) == "5-122"] <- "15-122"
colnames(counts_data)[colnames(counts_data) == "5-182"] <- "15-182"
colnames(counts_data)[colnames(counts_data) == "5-189"] <- "15-189"
colnames(counts_data)[colnames(counts_data) == "25-051"] <- "15-051"
colnames(counts_data)[colnames(counts_data) == "25-110"] <- "15-110"
colnames(counts_data)[colnames(counts_data) == "25-202"] <- "15-202"
colnames(counts_data)[colnames(counts_data) == "5-133"] <- "15-133"
head(counts_data) #shows the names and order of the rows. 
```

#Making vectors in counts.data matrix 
```{r}
sample_names <- colnames(counts_data) #gets the column names from counts_data and stores them in a vector called sample_names, making it easier to change and look at them.   
```

```{r}
sample_names #This shows me what the sample names are
```


```{r}
raw_groups <- sapply(strsplit(sample_names, "-"), `[`, 1) #splits the sample name wherever it finds a hyphen. "[, 1" says to take the first part of each split. "sapply" repeats the extraction for every sample leaving only the treatment number.
raw_groups
```

```{r}
groups <- factor(paste0("G", raw_groups)) #"paste0("G", raw_groups)" joins text together without spaces from the raw_groups. "G" is added in front of each number. Factor makes the labels "G#" into a category.
```

# making DGEList object

```{r}
y <- DGEList(counts = counts_data, group = groups) #function from the edgeR package that stores the count matrix, group labels, total reads per sample, normalization factors. "counts = counts_data" tells edgeR that they are the raw gene counts. "group = groups" tells edgeR which samples belong to which group. 
```

# filtering and normalization
* This filters out lowly expressed genes
```{r}
keep <- filterByExpr(y) #filters out genes with very low expression 
table(keep) # 48,616 genes were kept
y <- y[keep, , keep.lib.sizes=FALSE] #tells edgeR to get rid of the genes that are lowly expressed (FALSE). lib.sizes tells edgeR to update the library size based on the genes that are left in the dataset. 

y <- calcNormFactors(y) #normalizes the dataset so that it can be compared. 
```

# creating design matrix

```{r}
design <- model.matrix(~ 0 + groups) #creates a table that tells edgeR which samples belong to which experimental group. The "0" tells Rstudio to give each group its own column.   
colnames(design) <- levels(groups) #renames the columns to be the same as the group names "G5..." 
```

# fitting model

```{r}
y <- estimateDisp(y, design, robust=TRUE) #compares the difference in gene expression between the same samples.robust = TRUE keeps outliers from making a huge impact on gene expression differences.  

fit <- glmQLFit(y, design, robust=TRUE) #estimates how gene expression changes between the treatment groups while tracking biological variation.
```

# comparing 25C and 5C treatments 

```{r}
my.contrasts <- makeContrasts(Group25_vs_5 = G25 - G5, levels = design)
qlf_25_vs_5 <- glmQLFTest(fit, contrast=my.contrasts[,"Group25_vs_5"])
topTags(qlf_25_vs_5, n=10)
results_25vs5 <- topTags(qlf_25_vs_5, n=Inf)$table
filtered_25vs5 <- subset(results_25vs5, FDR < 0.05)
nrow(filtered_25vs5) # 33,733 DEGs 
# Comparing the expressed genes in treatments 25ºC and 5ºC. Displays the top 10 expressed genes as well as the ones that are significant (P<0.05)   
```

# comparing 15C and 5C treatments

```{r}
my.contrasts <- makeContrasts(Group15_vs_5 = G15 - G5, levels = design) #tells edgeR which groups you want to compare and puts the data under a new group called "my.contrasts"
qlf_15_vs_5 <- glmQLFTest(fit, contrast=my.contrasts[,"Group15_vs_5"]) #looks at which genes are expressed differently between 5 and 15ºC groups and htens tores the test results for each gene. 
topTags(qlf_15_vs_5, n=10) 
results_15vs5 <- topTags(qlf_15_vs_5, n=Inf)$table #makes the results from the previous line and formats them into a table 
filtered_15vs5 <- subset(results_15vs5, FDR < 0.05)
nrow(filtered_15vs5) # 8557 DEGs #filter through the results and only keep the ones that are statistically significant. 
```

# comparing 15C and 25C treatments

```{r}
my.contrasts <- makeContrasts(Group15_vs_25 = G15 - G25, levels = design)
qlf_15_vs_25 <- glmQLFTest(fit, contrast=my.contrasts[,"Group15_vs_25"]) #tells edgeR what values to contrast and what to label the new dataset.  
topTags(qlf_15_vs_25, n=10) #sorts genes by statistical significance (P<0.05)
results_15vs25 <- topTags(qlf_15_vs_25, n=Inf)$table #saves all the results that are significant in the table 
filtered_15vs25 <- subset(results_15vs25, FDR < 0.05) #filters out the results the have a P > 0.05
nrow(filtered_15vs25) # 888 DEGs #counts the # of rows in the new table 
```

# making the PCA (Principal Component Analysis)
```{r}
logCPM <- cpm(y, log = TRUE) #Converts the raw RNA-seq read counts into log Counts Per Million (logCPM) to normalize it.

pca_results <- prcomp(t(logCPM), scale. = TRUE) #Principal Component Analysis (PCA) which looks at all genes simultaneously and determines the combination of genes that show differences between samples.

pca_data <- data.frame(
  Sample = rownames(pca_results$x),
  PC1 = pca_results$x[, 1],
  PC2 = pca_results$x[, 2],
  Group = groups #creates a new table containing the information needed to make a PCA plot (2-D scatter plot of large datasets)
)

pca_var <- pca_results$sdev^2
pca_var_perc <- round(pca_var / sum(pca_var) * 100, 1) #calculates the % of transcription variance in the RNA-seq which helps with labeling the X-axis for the PC plot. 
```

# plotting the PCA 

```{r}
ggplot(pca_data, aes(PC1, PC2, group = Group, color = Group, shape = Group)) +
  geom_point(size=3) + #making a ggplot 
  stat_ellipse() + #Add 95% confidence interval ellipses around data
  xlab(paste0("PC1: ",pca_var_perc[1],"% variance")) + #Labels the x-axis with the PC1 % values.
  ylab(paste0("PC2: ",pca_var_perc[2],"% variance")) + #labels the y-axis with PC2 % values 
  coord_fixed() + #makes sure that distance on the plot is equal and the plot is not compressed or stretched.
  scale_color_manual(values = c( "#969696", "#CB181D", "#2171B5"), #color coordinator 
                     name = "Temperature (ºC)", #names the legend "temperature"
                     labels = c("15", "25", "5")) + #The different colors coordinate with the three temperatures. 
  scale_shape_manual(values = c(17, 16, 15), 
                     name = "Temperature (ºC)", 
                     labels = c("15", "25", "5")) + #picks different shapes for the three treatment data points. 
  theme_classic() + theme(legend.position = "top") #this makes the background for the plot white. 

ggsave("../output/06f-gene-expression/EdgeR_PCAcraball.pdf", width = 11, height = 8.5) #has Rstudio save the image in the designated location. 
```


