---
title: "06g-gene-expression-edgeR"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#loading required packages
```{r}
require(tidyverse)
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("DESeq2")
BiocManager::install("edgeR")
library(DESeq2)
library(ggplot2)
library(edgeR)
require("RColorBrewer")
install.packages("pheatmap")
require(pheatmap)
```
# link for the dataset download
```{r}
# https://osf.io/ea8mq/files/xnt56
```

# MAC import dataset
```{r}
require(tidyverse)
salmon.data <- read.delim("../../../Downloads/salmon.genes.counts.matrix", header=TRUE , check.names = FALSE)
```
# WINDOWS import dataset
```{r}
require(tidyverse)
salmon.data <- read.delim("../../../../Downloads/salmon.genes.counts.matrix", header=TRUE , check.names = FALSE)
```
# alter salmon.data so ID column doesn't mess with data
```{r}
colnames(salmon.data)[1] <- "ID"
head(salmon.data)
```

# making dataset renaming columns so they fit DGEList
```{r}
counts_data <- salmon.data
rownames(counts_data) <- counts_data[, 1]
counts_data <- counts_data[, -1]
sample_names <- colnames(counts_data)
raw_groups <- sapply(strsplit(sample_names, "-"), `[`, 1)
groups <- factor(paste0("G", raw_groups)) 
```

# making DGEList object

```{r}
y <- DGEList(counts = counts_data, group = groups)
```

# filtering and normalization

```{r}
keep <- filterByExpr(y)
table(keep) # 59,912 genes were kept
y <- y[keep, , keep.lib.sizes=FALSE]
y <- calcNormFactors(y)
```

# creating design matrix

```{r}
design <- model.matrix(~ 0 + groups)
colnames(design) <- levels(groups) 
```

# fitting model

```{r}
y <- estimateDisp(y, design, robust=TRUE)
fit <- glmQLFit(y, design, robust=TRUE)
```

# comparing 25C and 5C treatments

```{r}
my.contrasts <- makeContrasts(Group25_vs_5 = G25 - G5, levels = design)
qlf_25_vs_5 <- glmQLFTest(fit, contrast=my.contrasts[,"Group25_vs_5"])
topTags(qlf_25_vs_5, n=10)
results_25vs5 <- topTags(qlf_25_vs_5, n=Inf)$table
filtered_25vs5 <- subset(results_25vs5, FDR < 0.05)
nrow(filtered_25vs5) # 33,733 DEGs 
# change n= value to see more genes, or remove it to see all. Negative logFC means downregulation of first treatment vs second treatment
```

# comparing 15C and 5C treatments

```{r}
my.contrasts <- makeContrasts(Group15_vs_5 = G15 - G5, levels = design)
qlf_15_vs_5 <- glmQLFTest(fit, contrast=my.contrasts[,"Group15_vs_5"])
topTags(qlf_15_vs_5, n=10)
results_15vs5 <- topTags(qlf_15_vs_5, n=Inf)$table
filtered_15vs5 <- subset(results_15vs5, FDR < 0.05)
nrow(filtered_15vs5) # 8557 DEGs
```

# comparing 15C and 25C treatments

```{r}
my.contrasts <- makeContrasts(Group15_vs_25 = G15 - G25, levels = design)
qlf_15_vs_25 <- glmQLFTest(fit, contrast=my.contrasts[,"Group15_vs_25"])
topTags(qlf_15_vs_25, n=10)
results_15vs25 <- topTags(qlf_15_vs_25, n=Inf)$table
filtered_15vs25 <- subset(results_15vs25, FDR < 0.05)
nrow(filtered_15vs25) # 888 DEGs
```

# making the PCA
```{r}
logCPM <- cpm(y, log = TRUE)
pca_results <- prcomp(t(logCPM), scale. = TRUE)
pca_data <- data.frame(
  Sample = rownames(pca_results$x),
  PC1 = pca_results$x[, 1],
  PC2 = pca_results$x[, 2],
  Group = groups 
)

pca_var <- pca_results$sdev^2
pca_var_perc <- round(pca_var / sum(pca_var) * 100, 1)
```

# plotting the PCA 

```{r}
ggplot(pca_data, aes(x = PC1, y = PC2, color = Group)) +
  geom_point(size = 4, alpha = 0.8) +
  stat_ellipse(level = 0.95, linetype = 2) + 
  xlab(paste0("PC1: ", pca_var_perc[1], "% variance")) +
  ylab(paste0("PC2: ", pca_var_perc[2], "% variance")) +
  theme_minimal() + 
  ggtitle("PCA Plot of Treatment Groups") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    legend.title = element_text(face = "bold")
  )
```


